They're Clutching up! Team Momentum in Round-Based Esports

### 2022-08-13 paper/paper.qmd
trying paper with inline code and first plot

# 1. Introduction

## 1.1 Description of Call of Duty Search and Destroy

Call of Duty (CoD), first released in 2003, is one of the most popular first-person shooter (FPS) video game franchises of all-time. Tthe most popular mode in the competitive scene is "Search and Destroy" (SnD), which bears resemblance to "Bomb Defusal" in Counter-Strike and "Plant/Defuse" in Valorant, two other FPS games played in professional leagues. SnD is one-sided game mode in which one team, the offensive side, tries to destroy one of two designated bomb sites on the map.[^1]

[^1]: as

In professional CoD, a team must win six rounds of SnD to win the match.[^2] A round can end in one of four ways:

[^2]: A maximum of 11 even rounds can be played. There is no "sudden death" or "win by two" rule like there are for SnD equivalent in professional Counter-Strike and Valorant matches.

1.  One team eliminates all members of the other team prior to a bomb plant. (Eliminating team wins.)
2.  The offensive team eliminates all members of the defensive team after a bomb plant.[^3] (Offense wins.)
3.  The defensive team defuses the bomb after a bomb plant.[^4] (Defense wins.)
4.  The offensive team does not make a plant by the time the round timer ends. (Defense wins.)

[^3]:
    -   The bomb can be picked up by any member of the offensive team.
    -   The bomb carrier is not obstructed at all by carrying the bomb (i.e. movement is the same, weapon usage is the same).
    -   The defense does not get any visual indication for who is carrying the bomb.
    -   A bomb plant takes five seconds. The timer resets if the player stops planting site prior to completing it.
    -   A bomb defuse takes seven seconds. The timer resets if the player "drops" the bomb.
    -   The bomb takes 45 seconds to defuse after being planted.

[^4]: Often the defensive team will try to eliminate all team members prior to making the defuse, but in some cases, they may try to "ninja" defuse.

@@

## 1.2 Data

CoD has roughly gone through three eras of professional gaming: (1) Major League Gaming (MLG) tournaments prior to 2016; (2) the CoD World League (CWL), initiated in 2016; and (3) the 12-franchise CoD League (CDL), running since 2020 and completing three year-long "seasons" completed as of August 2022.[^5] The data set consists of all SnD matches played in tournaments and qualifiers during the CDL era. Data was collected in spreadsheets by community member "IOUTurtle".[^6]

[^5]: CoD is fairly unique compared to other esports in that it runs on an annual lifecycle (released coming in the late fall), where a new game is published every year under the same title. Each new game bears resemblance to past ones, often introducing relatively small variations ("improvements") to graphics, game modes, and other facets of gameplay. During the CDL era, the games released have been Modern Warfare (2020), Cold War (2021) and Vanguard (2022).

[^6]: Data: https://linktr.ee/CDLArchive. Author: https://twitter.com/IOUTurtle

@@

## 1.3 My contribution

# 2 Literature review

@@

Despite the plethora of existing research on games played and momentum in sports, these topics have yet to be investigated heavily in esports. Work has been done to examine in-round win probability in other FPS titles such as Counter-Strike ([@xenopoulos2022]) and Valorant ([@derover2021]), both of which are round-based like CoD SnD. However, research on round-level trends is sparse, perhaps because games like Counter-Strike and Valorant both have economic aspects that can create clear advantages on side in a given round, given how prior rounds played out.[^7]\[\^8\]

[^7]: Additionally, both Counter-Strike and Valorant have overtime rules and blocked offensive/defensive roles (i.e. playing either offense or defense for many consecutive rounds).

# 3. Methodology

# 4. Results and discussion

>>>>

```{r}
#| label: setup
#| eval: true
library(dplyr)
library(dplyr)
library(qs)
library(scales)

cod_rounds <- qs::qread('cod_rounds.qs')
```

```{r}
#| label: descriptive-cod-snd-stats
#| eval: true
cod_n_series <- cod_rounds |> 
  distinct(series_id, team) |> 
  nrow()

cod_n_rounds <- cod_rounds |> 
  distinct(series_id, team, round) |> 
  nrow()

cod_o_win_prop <- cod_rounds |> 
  filter(is_offense) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop)

cod_o_win_prop_by_game <- cod_rounds |> 
  filter(is_offense) |> 
  count(game, win_round) |> 
  group_by(game) |> 
  mutate(prop = n / sum(n)) |> 
  ungroup() |> 
  filter(win_round) |> 
  select(game, n, prop)

pull_cod_o_win_prop_by_game <- function(.game) {
  cod_o_win_prop_by_game |> 
    filter(game == .game) |> 
    pull(prop) |> 
    scales::percent(accuracy = 0.1)
}
```

# Introduction

## 1 Description of Call of Duty Search and Destroy

Call of Duty (CoD), first released in 2003, is one of the most popular first-person shooter (FPS) video game franchises of all-time. Tthe most popular mode in the competitive scene is "Search and Destroy" (SnD), which bears resemblance to "Bomb Defusal" in Counter-Strike and "Plant/Defuse" in Valorant, two other FPS games played in professional leagues. SnD is one-sided game mode in which one team, the offensive side, tries to destroy one of two designated bomb sites on the map.

In professional CoD, a team must win six rounds of SnD to win the match.[^1] A round can end in one of four ways:

[^1]: A maximum of 11 even rounds can be played. There is no "sudden death" or "win by two" rule like there are for SnD equivalent in professional Counter-Strike and Valorant matches.

1.  One team eliminates all members of the other team prior to a bomb plant. (Eliminating team wins.)
2.  The offensive team eliminates all members of the defensive team after a bomb plant.[^2] (Offense wins.)
3.  The defensive team defuses the bomb after a bomb plant.[^3] (Defense wins.)
4.  The offensive team does not make a plant by the time the round timer ends. (Defense wins.)

[^2]:
    -   The bomb can be picked up by any member of the offensive team.
    -   The bomb carrier is not obstructed at all by carrying the bomb (i.e. movement is the same, weapon usage is the same).
    -   The defense does not get any visual indication for who is carrying the bomb.
    -   A bomb plant takes five seconds. The timer resets if the player stops planting site prior to completing it.
    -   A bomb defuse takes seven seconds. The timer resets if the player "drops" the bomb.
    -   The bomb takes 45 seconds to defuse after being planted.

[^3]: Often the defensive team will try to eliminate all team members prior to making the defuse, but in some cases, they may try to "ninja" defuse.

@@

## 1. Data

CoD has roughly gone through three eras of professional gaming: (1) Major League Gaming (MLG) tournaments prior to 2016; (2) the CoD World League (CWL), initiated in 2016; and (3) the 12-franchise CoD League (CDL), running since 2020 and completing three year-long "seasons" completed as of August 2022.[^4] The data set consists of all SnD matches played in tournaments and qualifiers during the CDL era, totaling `r scales::comma(cod_n_series)` series played, in which there were `r scales::comma(cod_n_rounds)` rounds played. Data was collected in spreadsheets by community member "IOUTurtle".[^5]

[^4]: CoD is fairly unique compared to other esports in that it runs on an annual lifecycle (released coming in the late fall), where a new game is published every year under the same title. Each new game bears resemblance to past ones, often introducing relatively small variations ("improvements") to graphics, game modes, and other facets of gameplay. During the CDL era, the games released have been Modern Warfare (2020), Cold War (2021) and Vanguard (2022).

[^5]: Data: https://linktr.ee/CDLArchive. Author: https://twitter.com/IOUTurtle

@@

# Literature review

@@

Despite the plethora of existing research on games played and momentum in sports, these topics have yet to be investigated heavily in esports. Work has been done to examine in-round win probability in other FPS titles such as Counter-Strike ([@xenopoulos2022]) and Valorant ([@derover2021]), both of which are round-based like CoD SnD. However, research on round-level trends is sparse, perhaps because games like Counter-Strike and Valorant both have economic aspects that can create clear advantages on side in a given round, given how prior rounds played out.[^6]

[^6]: Additionally, both Counter-Strike and Valorant have overtime rules and blocked offensive/defensive roles (i.e. playing either offense or defense for many consecutive rounds).

# Methodology, results, and discussion

The empirical offensive round win percentage across all rounds is `r scales::percent(cod_o_win_prop, accuracy = 0.1)`.[^7] @fig-cod-o-round-win-prop shows that offensive round win rate is not quite constant, although, on average, never veers more than 10% from this global average.

[^7]: Offensive round win percentage has been nearly constant across the three games during the CDL era: MW (2020) `r pull_cod_o_win_prop_by_game('MW (2020)')`, Cold War (2021) `r pull_cod_o_win_prop_by_game('Cold War (2021)')`, Vanguard (2022) `r pull_cod_o_win_prop_by_game('Vanguard (2022)')`

::: {#fig-cod-o-round-win-prop}
![](images/cod_o_round_win_prop.png){fig-align="center"}

Offensive round win percentage by series state
:::

### 2022-08-13 paper/paper.tex
trying paper with inline code and first plot

Finally, I evaluate behavior when teams have two rounds left to win the
series, observing a peak in COD offensive win percentages in the 4-4
state, and no such oddity in Valorant.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{1. Introduction}\label{introduction}}

\hypertarget{description-of-call-of-duty-search-and-destroy}{%
\subsection{1.1 Description of Call of Duty Search and
Destroy}\label{description-of-call-of-duty-search-and-destroy}}

Call of Duty (CoD), first released in 2003, is one of the most popular
first-person shooter (FPS) video game franchises of all-time. Tthe most
popular mode in the competitive scene is ``Search and Destroy'' (SnD),
which bears resemblance to ``Bomb Defusal'' in Counter-Strike and
``Plant/Defuse'' in Valorant, two other FPS games played in professional
leagues. SnD is one-sided game mode in which one team, the offensive
side, tries to destroy one of two designated bomb sites on the
map.\footnote{as}

@@

\hypertarget{data}{%
\subsection{1.2 Data}\label{data}}

CoD has roughly gone through three eras of professional gaming: (1)
Major League Gaming (MLG) tournaments prior to 2016; (2) the CoD World
League (CWL), initiated in 2016; and (3) the 12-franchise CoD League
(CDL), running since 2020 and completing three year-long ``seasons''
completed as of August 2022.\footnote{CoD is fairly unique compared to
  other esports in that it runs on an annual lifecycle (released coming
  in the late fall), where a new game is published every year under the
  same title. Each new game bears resemblance to past ones, often
  introducing relatively small variations (``improvements'') to
  graphics, game modes, and other facets of gameplay. During the CDL
  era, the games released have been Modern Warfare (2020), Cold War
  (2021) and Vanguard (2022).} The data set consists of all SnD matches
played in tournaments and qualifiers during the CDL era. Data was
collected in spreadsheets by community member ``IOUTurtle''.\footnote{Data:
  https://linktr.ee/CDLArchive. Author: https://twitter.com/IOUTurtle}

@@

\hypertarget{my-contribution}{%
\subsection{1.3 My contribution}\label{my-contribution}}

\hypertarget{literature-review}{%
\section{2 Literature review}\label{literature-review}}

@@

Despite the plethora of existing research on games played and momentum
in sports, these topics have yet to be investigated heavily in esports.
Work has been done to examine in-round win probability in other FPS
titles such as Counter-Strike ((Xenopoulos, Freeman, and Silva 2022))
and Valorant ((DeRover 2021)), both of which are round-based like CoD
SnD. However, research on round-level trends is spare, perhaps because

>>>>

Finally, I evaluate behavior when teams have two rounds left to win the
series, observing a peak in COD offensive win percentages in the 4-4
state, and no such oddity in Valorant.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, boxrule=0pt, breakable, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{description-of-call-of-duty-search-and-destroy}{%
\subsection{1 Description of Call of Duty Search and
Destroy}\label{description-of-call-of-duty-search-and-destroy}}

Call of Duty (CoD), first released in 2003, is one of the most popular
first-person shooter (FPS) video game franchises of all-time. Tthe most
popular mode in the competitive scene is ``Search and Destroy'' (SnD),
which bears resemblance to ``Bomb Defusal'' in Counter-Strike and
``Plant/Defuse'' in Valorant, two other FPS games played in professional
leagues. SnD is one-sided game mode in which one team, the offensive
side, tries to destroy one of two designated bomb sites on the map.

@@

\hypertarget{data}{%
\subsection{1. Data}\label{data}}

CoD has roughly gone through three eras of professional gaming: (1)
Major League Gaming (MLG) tournaments prior to 2016; (2) the CoD World
League (CWL), initiated in 2016; and (3) the 12-franchise CoD League
(CDL), running since 2020 and completing three year-long ``seasons''
completed as of August 2022.\footnote{CoD is fairly unique compared to
  other esports in that it runs on an annual lifecycle (released coming
  in the late fall), where a new game is published every year under the
  same title. Each new game bears resemblance to past ones, often
  introducing relatively small variations (``improvements'') to
  graphics, game modes, and other facets of gameplay. During the CDL
  era, the games released have been Modern Warfare (2020), Cold War
  (2021) and Vanguard (2022).} The data set consists of all SnD matches
played in tournaments and qualifiers during the CDL era, totaling 1,704
series played, in which there were 15,584 rounds played. Data was
collected in spreadsheets by community member ``IOUTurtle''.\footnote{Data:
  https://linktr.ee/CDLArchive. Author: https://twitter.com/IOUTurtle}

@@

\hypertarget{literature-review}{%
\section{Literature review}\label{literature-review}}

@@

Despite the plethora of existing research on games played and momentum
in sports, these topics have yet to be investigated heavily in esports.
Work has been done to examine in-round win probability in other FPS
titles such as Counter-Strike ((Xenopoulos, Freeman, and Silva 2022))
and Valorant ((DeRover 2021)), both of which are round-based like CoD
SnD. However, research on round-level trends is sparse, perhaps because
games like Counter-Strike and Valorant both have economic aspects that
can create clear advantages on side in a given round, given how prior
rounds played out.\footnote{Additionally, both Counter-Strike and
  Valorant have overtime rules and blocked offensive/defensive roles
  (i.e.~playing either offense or defense for many consecutive rounds).}

\hypertarget{methodology-results-and-discussion}{%
\section{Methodology, results, and
discussion}\label{methodology-results-and-discussion}}

The empirical offensive round win percentage across all rounds is
47.8\%.\footnote{Offensive round win percentage has been nearly constant
  across the three games during the CDL era: MW (2020) 47.2\%, Cold War
  (2021) 47.9\%, Vanguard (2022) 48.1\%}
Figure~\ref{fig-cod-o-round-win-prop} shows that offensive round win
rate is not quite constant, although, on average, never veers more than
10\% from this global average.

\begin{figure}

{\centering

\begin{figure}

{\centering \includegraphics{images/cod_o_round_win_prop.png}

}

\end{figure}

}

\caption{\label{fig-cod-o-round-win-prop}Offensive round win percentage
by series state}

\end{figure}
